---
title: "Select and plot bromeliaceae distribution and diversity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 8,
  collapse = TRUE,
  comment = "#>"
)
```

The bromeliad package contains estimated distributions for 2395 species of the pineapple family (Bromeliaceae) plus three functions to access and visualize this data: `get_range`, `map_species`, `map_richness`. The included ranges were estimated from occurrence records, using different algorithms, including climate suitability models based on climate, convex hulls and geospatial buffering. Make sure to check Zizka *et al.* 2019 for the methodology of range estimation. 

# Libraries and installation
You can install bromeliad using the devtools package. If you want to visualize species richness patterns, you also need to install speciesgeocodeR.

```{r, eval = T}
library(devtools)

#install_github("idiv-biodiversity/bromeliad")
#install_github("azizka/speciesgeocodeR")

library(bromeliad)
library(speciesgeocodeR)
```


# Selection and vizualisation

If you want to use data for all bromeliaceae, you can access them directly via the dataset using `data()`. If you are only interested in a subset of the species you can use the `get_range` function to select species based on taxonomy (scientific name, canonical name, genus, subfamily) or traits (life form, range size, conservation assessment), or a combination of those.

You can select one or multiple species and visualize their range as per species map. The type argument controls what kind of range estimation is selected.

```{r}
# range polygon based on scientific name with authorities
dat <- get_range(scientific = "Aechmea mexicana Baker")

# presence absence model
dat <- get_range(scientific = "Aechmea alba Mezr", type = "binary")

#modelled habitat suitability
dat <-  get_range(scientific = "Aechmea alba Mez", type = "suitability")

# based on canonical name
dat <- get_range(canonical = "Aechmea mexicana")

# multiple species
dat <- get_range(canonical = c("Aechmea mexicana", "Aechmea alba"))

# Species maps
map_species(canonical = "Aechmea mexicana")

# Richness map 
map_richness(dat)
```

You can also use higher rank taxa to select multiple species at once.
```{r}
# one genus
dat <-  get_range(genus = "Aechmea")

# multiple genera
dat <- get_range(genus = c("Aechmea", "Zizkaea"))

# subfamily
dat <- get_range(subfamily = "Pitcairnioideae")

#map_species(dat) This will generate many maps! You can send them to a pdf:
# pdf("achmea_ranges.pdf")
# map_species(genus = "Aechmea")
# dev.off()

#Richness pattern
map_richness(dat)
```


Or based on traits and conservation status.

```{r}
#life form
dat <- get_range(life_form = "epiphyte")
map_richness(dat)

# conservation status
dat <-  get_range(assessment = c("CR", "VU"))
```

Different criteria can also be combined, for instance to get all species of *Aechmea* classified as critcally endangered, endangered or vulnerable.

```{r}
dat <-  get_range(genus = "Aechmea", assessment = c("CR", "EN", "VU"))
map_richness(dat)
```


